# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.
concise_data_ncd <- data_ncd %>%
  rename(DALYs = "DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)")


subset_wider <- concise_data_ncd %>%
  filter(Entity %in% c("Mali", "England", "Maldives")) %>% 
  pivot_wider(names_from = Entity,
              values_from = DALYs, id_cols = Year)
kable(subset_wider)
```

## Plot Showing Trends in NCD Burden Over Time 

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
selected_countries <- concise_data_ncd %>% 
  filter(Entity %in% c("Mali", "England", "Maldives"))

DALYs_trend_line <- ggplot(data = selected_countries,
                          aes(x= Year, y = DALYs, group = Entity, color = Entity)) +
  geom_line() +
  labs(title = "NCD Burden over time in selected countries",
       x = "Year",
       y = "DALYs lost") +
  theme_minimal()
DALYs_trend_line
```

## Summary of NCD Burden Findings

NCD burden changed the most dramatically in the Maldives where it halved during the period of 30 years. DALYs lost in England and Mali shared a similar trend in reduction but with Mali having more a higher burden.

